---
title: "194.044 Data Stewardship"
subtitle: "Assignment 1"
author: "Aleksandar Pavlovic"
date: "16.04.2019"
output: pdf_document
---

# Question
This document deals with answering the pseudo question, if the global temperature trend influences the number of children born alive in Vienna.

# Datasets
To answer the proposed question above, the following two datasets are used:
\begin{itemize}
\item
Live births in Vienna - identifier: f54e6828-3d75-4a82-89cb-23c58057bad4, url: [birth data repo](https://www.data.gv.at/katalog/dataset/f54e6828-3d75-4a82-89cb-23c58057bad4)
\item
Mean near surface temperature deviation (the first zip file was downloaded) - identifier: $cli\_iad\_td$, url: [temperature repo](https://data.europa.eu/euodp/en/data/dataset/zQAEvhkR7H0NQYU1HP5fA)
\end{itemize}

```{r}
# Load libraries
library("ggplot2")
```

```{r}
# Load preprocessed and merged data
data <- read.csv2("./data/processed/mergedData.csv")
```

```{r}
# First visualisation:
ggplot(data=data, aes(x=year, group=1)) +
  labs(y = "Global mean temperature deviation",
                x = "Years") +
  geom_line(aes(y = GlobalTempDev, colour = "Temp dev")) +
  geom_line(aes(y = LIVEBIRTH/20000, colour = "Live-births")) +
  scale_y_continuous(sec.axis = sec_axis(~.*20000, name = "Number of live-births"))

ggsave("./plots/timeSeriesLinePlot.jpg")
```

One can see in the graph above, that the general trend of both variables is the same (thus the two variables seem to be highly correlated).

```{r}
# Second visualisation:
ggplot(data=data, aes(x=GlobalTempDev, group=1)) +
  geom_point(aes(y = LIVEBIRTH)) +
  geom_smooth(mapping = aes(x = GlobalTempDev, y = LIVEBIRTH), method=lm)

ggsave("./plots/scatterPlotWithLinearModel.jpg")
```
Further investigation leads to the above scatterplot enhanced with the prediction of a linear model (blue line) and the 95\% confidence interval (grey area). One can see, that all black points (data points) are very close to that prediction line, thus again revealing, that the two variables are highly correlated. 

```{r}
cor(data$GlobalTempDev, data$LIVEBIRTH, method="pearson")
```
Looking at the Pearson correlation coefficient reveals, that the data is in fact really highly correlated, thus when one ignores that there is a difference between causality and correlation, one can infer that the rising temperature leads to an increase of live-births in Vienna.
